\documentclass{beamer}
\usetheme{CambridgeUS}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,fit,positioning, mindmap, trees, automata}

\usepackage[latin1]{inputenc}
\usefonttheme{professionalfonts}
\usepackage{times}
\usepackage{xmpmulti}
\usepackage{animate}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{mathrsfs}  
\usepackage{bussproofs}

\usepackage{stmaryrd}
\usepackage{xcolor}

\newcommand{\person}[1]{\textcolor{blue}{#1}}
\newcommand{\highlight}[1]{\textcolor{red}{#1}}
% \newcommand{\example}[1]{\textcolor{blue}{#1}}
\usetikzlibrary{shapes.geometric, positioning}
\graphicspath{ {./images/} }
%\usetheme{Boadilla}
%\usecolortheme{crane}
\title[Computational Theory and Complexity Theory]{Computational Theory and Complexity Theory}
%\subtitle{I Am Curious}
\author[Sebastian Schlesinger]{Prof. Dr.-Ing. Sebastian Schlesinger}
\institute[HWR Berlin]{Berlin School for Economics and Law}
\date{\today}
\begin{document}
 \begin{frame}
\titlepage
\end{frame}



\begin{frame}{Word Problem}
\begin{itemize}
    \item Given a formal language $L$ over an alphabet $\Sigma$ and a word $w \in \Sigma^*$
    \item Decide whether $w \in L$
    \item Example: $L = \{a^n b^n | n \geq 0\}$, $w = aaabbb$
    \item $w \in L$?
\end{itemize}
\end{frame} 

\begin{frame}{Decidability and Semidecidability}
\begin{itemize}
    \item A language $L$ is \textbf{decidable} if there exists a Turing machine that halts on every input and accepts exactly the words in $L$
    \item A language $L$ is \textbf{semidecidable} if there exists a Turing machine that accepts exactly the words in $L$ and may either reject or run forever on words not in $L$
    \item Every decidable language is also semidecidable
    \item Example: The Halting Problem is semidecidable but not decidable
\end{itemize}
\end{frame}

\begin{frame}{Computability}
\begin{itemize}
    \item A function $f: \Sigma^* \rightarrow \Sigma^*$ is \textbf{computable} if there exists a Turing machine that, given any input $w \in \Sigma^*$, halts and outputs $f(w)$
    \item Example: The function that maps a binary number to its successor is computable
    \item Not all functions are computable (e.g., the Halting Problem)
\end{itemize}
\end{frame}

\begin{frame}{Completeness and Decidability}
 The propositional logic is decidable. However, first-order logic is not decidable, but semi-decidable.
 However, first-order logic is complete, i.e., every valid formula can be proven.
 Hence, although there must be a proof for every valid formula, there is no algorithm that can find it for every formula in finite time.
\end{frame}

\begin{frame}{Computability vs. Complexity}
\begin{itemize}
    \item \textbf{Computability} is concerned with what can be computed (i.e., the existence of an algorithm).
    \item \textbf{Complexity} is concerned with how efficiently a problem can be solved (i.e., the resources required).

\end{itemize}
\end{frame}

\begin{frame}{Complexity Classes}
\begin{itemize}
    \item \textbf{P}: Class of decision problems solvable by a deterministic Turing machine in polynomial time.
    \item \textbf{NP}: Class of decision problems solvable by a nondeterministic Turing machine in polynomial time. Equivalently, problems for which a given solution can be verified in polynomial time.
    \item \textbf{PSPACE}: Class of decision problems solvable by a Turing machine using polynomial space.
    \item \textbf{EXPTIME}: Class of decision problems solvable by a deterministic Turing machine in exponential time.
\end{itemize}
\end{frame} 
\begin{frame}{P vs NP Problem}
\begin{itemize}
    \item One of the most important open problems in computer science.
    \item Question: Is P equal to NP?
    \item If P = NP, then every problem for which a solution can be verified quickly can also be solved quickly.
    \item Most experts believe that P $\neq$ NP, but it remains unproven.
\end{itemize}
\end{frame} 
\begin{frame}{NP-Hard and NP-Complete}
\begin{itemize}
    \item A problem is \textbf{NP-hard} if every problem in NP can be reduced to it in polynomial time.
    \item A problem is \textbf{NP-complete} if it is both in NP and NP-hard.
    \item If any NP-complete problem can be solved in polynomial time, then P = NP.
    \item Examples of NP-complete problems: Traveling Salesman Problem, Boolean Satisfiability Problem (SAT), Knapsack Problem
\end{itemize}
\end{frame}
\end{document}